%--------------------------------------------------------------------------
% Acknowledgements
%--------------------------------------------------------------------------

This thesis is the product of perseverance and hard work from my long, meandering journey in pursuit of achieving a master’s degree, and of wanting to do research that not only aligned with my own interests in applying remote sensing for forest conservation and management, but also of contributing an academic piece of work with a practical purpose. It is my hope that this thesis is found in the hands of individuals who by reading, studying, critiquing, and improving it find value in this work\textemdash all to further contribute to the greater body of knowledge\textemdash rather than just ending up unread in shelves, gathering dust.
Completing this master’s thesis would not have been possible without the support and encouragement of family, friends, and colleagues, especially when my commitment to family and professional work took precedence over finishing my postgraduate studies. I am so grateful to all of you.
To Dr Enrico Paringit, my thesis adviser and professor, for sharing his knowledge and providing guidance in doing my thesis, and for his patience, understanding, persistence, and encouragement all throughout my master’s studies.
To Dr Rijaldia Santos and Engr Ma Rosario Concepcion Ang, both of whom were my former professors during my undergraduate and masters studies, respectively. Thank you for agreeing to be my thesis panel members and providing very helpful comments during my thesis viva and for improving the manuscript.
To this extraordinary group of friends, colleagues, research group mates wherein our professional, academic, and even personal lives have been intertwined: Angelica Kristina Monzon, Rizza Karen Veridiano, Margie Parinas, Sheryl Rose Reyes, Roven Tumaneng, Ray Neil Lorca, Raul Magabo, and Francisco Tavora Jr\textemdash this journey was all the more worthwhile because we were all in the thick of things together.
To my former professors, including Dr Ariel Blanco, Dr Rhodora Gonzalez, and Dr Ma Cecilia Rubio-Paringit, who kept encouraging me to finish my thesis at every opportunity that I meet them. This also goes for my boss at my new research work at the National University of Singapore, Dr Edward Webb\textemdash thanks for keeping me on my toes.
To close friends and colleagues: Edmund Leo Rico, Ronald Allan Altamirano, and Dr Neil Aldrin Mallari\textemdash thank you all for those long discussions, brainstorming sessions, and exchange of ideas of the things that make us look forward to waking up in the morning.
I am also grateful to Mari Trix Estomata, Nimol Vameourn, Christian Supsup, Dennis Tacmo, Dr Edwino Fernando, and Ms Rhudelene Serrano who have all supported me in small ways but nevertheless added up to help me put my thesis together.
This work has been undertaken within the framework of the Kyoto \& Carbon Initiative led by Japan Aerospace Exploration Agency. The ALOS/PALSAR data used in this study have been provided by the JAXA Earth Observation Research Center.
I dedicate this academic work to Rina, my wife and best friend, and my twin daughters, Dahon Luntian and Ulap Amihan\textemdash you are all my inspiration.
And to God for whom all of this was meant for his greater honor and glory.